---
title: 'Setup, Quality Control and Sample Integration.'
teaching: 10
exercises: 2
---

:::::::::::::::::::::::::::::::::::::: questions 

- How do you write a lesson using R Markdown and `{sandpaper}`?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Explain how to use markdown with the new lesson template
- Demonstrate how to include pieces of code, figures, and nested challenge blocks

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: callout

This tutorial is partially based on existing material from:

- https://satijalab.org/seurat/articles/seurat5_integration
- https://satijalab.org/seurat/articles/de_vignette
- https://hbctraining.github.io/scRNA-seq_online/
- https://bookdown.org/ytliu13207/SingleCellMultiOmicsDataAnalysis/

:::::::::::::::::::::::::::::::::::::::::::::


## Setup
**Run the code block below to install the packages needed for this workshop.**

```{r install, results='hide', message=F, warning=F}
# install.packages('Seurat')
# if (!requireNamespace("remotes", quietly = TRUE)) {
#   install.packages("remotes")
# }
# remotes::install_github("satijalab/seurat-data", quiet = TRUE)
# if (!require("BiocManager", quietly = TRUE)) {
#   install.packages("BiocManager")
# }
# BiocManager::install("DESeq2")
# install.packages("tidyverse")
# install.packages("pheatmap")
# install.packages("metap")
```

To check if installed properly, load each package in one at a time using the `library()` function.

```{r setup, results='hide', message=F, warning=F}
library(Seurat, warn.conflicts=F, quietly=T)
#library(SeuratData, warn.conflicts=F, quietly=T)
library(tidyverse, warn.conflicts=F, quietly=T)
library(DESeq2, warn.conflicts=F, quietly=T)
library(patchwork, warn.conflicts=F, quietly=T)
library(pheatmap, warn.conflicts=F, quietly=T)
library(grid, warn.conflicts=F, quietly=T)
library(metap, warn.conflicts=F, quietly=T)

set.seed(4242) # Set Seed for Reproducibility
```

We're using the ifnb public dataset provided by Seurat. This dataset contains PBMC data from 8 lupus patients before and after interferon beta treatment.

I strongly encourage you to explore the other datasets offered by the SeuratData package, it can be really good practice in your spare time.

The ifnb Seurat object we're loading in here was originally made in Seurat v4, there have since been a lot of changes from Seurat v4 to v5 so we'll use the UpdateSeuratObject() function to update the Seurat object so that it is compatible for today.

!!! question
    Looking at the output from the `str()` function on the ifnb seurat object, can you tell whether this seurat object is processed or unprocessed?

ANSWER: When loading in seurat objects, we can have a look at what processing steps have been performed on it by using the str() function. In the output we can tell that the ifnb Seurat object is unprocessed because:

1)  the scale.data slot is empty

2)  no variable features have been identified

3)  no dimensionality reduction functions have been performed.

```{r available-data}
# AvailableData() # if you want to see the available datasets use this function
```
